---
title: "Homework for Network Analysis Workshop"
author: "Min Hee Seo"
date: "August 21st, 2018"
output: pdf_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

\section{Exercise 1. Nigeria Data}

```{r ex1, eval=T, echo=T, warning=F, message=F, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
# create network matrix by year and all years

# setup the working directory
setwd("/Users/minheeseo/Dropbox/Classes/2018_Classes/Network/network2018_hw1/")
# clean the workspace
rm(list=ls())
# loading data and R packages
library(igraph)
library(network)
load("nigeria.rda")
# clean the labels
nigeria$sender<- gsub("\n", " ", nigeria$sender)
nigeria$receiver <- gsub("\n", " ", nigeria$receiver)

# create an empty list where the length of list is time span
network.mat <- vector("list",length(unique(nigeria$year)))
names(network.mat) <- unique(nigeria$year)
# For each year,
# this loop stores one if there was a conflict between sender and receiver
# it stores 0 otherwise
# outcome, network.mat, is a list which contains 17 matrix for each year
time <- unique(nigeria$year)
country.sender <- unique(nigeria$sender)
country.receiver <- unique(nigeria$receiver)
for(t in 1:length(time)){
  slice <- NULL
  empty.mat <- NULL
  empty.mat <-matrix(0, length(country.sender), length(country.receiver))
  empty.mat <- as.data.frame(empty.mat)
  rownames(empty.mat) <- country.sender
  colnames(empty.mat) <- country.receiver
  slice <- subset(nigeria, nigeria$conflict==1 & nigeria$year==time[t])
  for(i in 1:nrow(slice)){
    empty.mat[rownames(empty.mat)==slice$sender[i], colnames(empty.mat)==slice$receiver[i]] <- 1
  }
  network.mat[[t]] <-empty.mat
}

# collapse each matrix into one
tmp <- Reduce("+", network.mat)
# add to an existing list
network.mat[[18]] <- tmp
names(network.mat)[18] <- "All Conflict"
```
After I created 17 matrices for each year and another one for all years, I plotted them.

```{r ex1plot, eval=T, echo=T, warning=F, message=F}
# plot network by each year and all years

myblue <- rgb(red = 0, green = 0, blue = 1, alpha = .5)
mypink <- rgb(red = 1, green = 0, blue = 0, alpha = .25)

par(mfrow=c(2, 2), mar=c(0,0.2,1,0.2))
for(i in 1:4){
  g <- NULL
  g = graph_from_adjacency_matrix(as.matrix(network.mat[[i]]),
                                mode='directed', weighted=TRUE, diag=F)
  tiesSum = apply(g[], 1, sum)
# condition size based on # of ties
  V(g)$size <- (tiesSum+0.5)*6
# only label if # ties greater than 10
  V(g)$label <- ifelse( tiesSum>1, V(g)$name, NA )
  V(g)$label.cex <- 0.6
  plot(g,main=paste("Year:", names(network.mat)[i]),
     vertex.label=V(g)$label, 
     vertex.size=V(g)$size,
     edge.width=E(g)$weight,
     vertex.color =mypink, # change color of nodes
     vertex.label.color = "black", # change color of labels
     edge.curved=.25, # add a 25% curve to the edges
     edge.color=myblue, # change edge color to grey  
     layout=layout_with_fr,
     edge.arrow.size=0.2)
}

```


```{r ex1plot2, eval=T, echo=F, warning=F, message=F}

par(mfrow=c(2, 2), mar=c(0,0.2,1,0.2))
for(i in 5:8){
  g <- NULL
  g = graph_from_adjacency_matrix(as.matrix(network.mat[[i]]),
                                mode='directed', weighted=TRUE,diag=F)
  tiesSum = apply(g[], 1, sum)
# condition size based on # of ties
  V(g)$size <- (tiesSum+0.5)*6
# only label if # ties greater than 10
  V(g)$label <- ifelse( tiesSum>1, V(g)$name, NA )
  V(g)$label.cex <- 0.6
  plot(g,main=paste("Year:", names(network.mat)[i]),
     vertex.label=V(g)$label, 
     vertex.size=V(g)$size,
     edge.width=E(g)$weight,
     vertex.color =mypink, # change color of nodes
     vertex.label.color = "black", # change color of labels
     edge.curved=.25, # add a 25% curve to the edges
     edge.color=myblue, # change edge color to grey  
     layout=layout_with_fr,
     edge.arrow.size=0.2)
}
```

```{r ex1plot3, eval=T, echo=F, warning=F, message=F}
par(mfrow=c(2, 2), mar=c(0,0.2,1,0.2))
for(i in 9:12){
  g <- NULL
  g = graph_from_adjacency_matrix(as.matrix(network.mat[[i]]),
                                mode='directed', weighted=TRUE,diag=F)
  tiesSum = apply(g[], 1, sum)
# condition size based on # of ties
  V(g)$size <- (tiesSum+0.5)*6
# only label if # ties greater than 10
  V(g)$label <- ifelse( tiesSum>1, V(g)$name, NA )
  V(g)$label.cex <- 0.6
  plot(g,main=paste("Year:", names(network.mat)[i]),
     vertex.label=V(g)$label, 
     vertex.size=V(g)$size,
     vertex.color=mypink,
     edge.width=E(g)$weight,
     vertex.label.color="black", 
     edge.curved=.25, 
     edge.color=myblue,
     layout=layout_with_fr,
     edge.arrow.size=0.2)
}
```

```{r ex1plot4, eval=T, echo=F, warning=F, message=F}
par(mfrow=c(2, 2), mar=c(0,0.2,1,0.2))
for(i in 13:16){
  g <- NULL
  g = graph_from_adjacency_matrix(as.matrix(network.mat[[i]]),
                                mode='directed', weighted=TRUE,diag=F)
  tiesSum = apply(g[], 1, sum)
# condition size based on # of ties
  V(g)$size <- (tiesSum+0.5)*6
# only label if # ties greater than 10
  V(g)$label <- ifelse( tiesSum>1, V(g)$name, NA )
  V(g)$label.cex <- 0.6
  plot(g,main=paste("Year:", names(network.mat)[i]),
     vertex.label=V(g)$label, 
     vertex.size=V(g)$size,
     edge.width=E(g)$weight,
     vertex.color=mypink, # change color of nodes
     vertex.label.color="black", # change color of labels
     edge.curved=.25, # add a 25% curve to the edges
     edge.color=myblue, # change edge color to grey  
     layout=layout_with_fr,
     edge.arrow.size=0.2)
}

```


```{r ex1plot5, eval=T, echo=F, warning=F, message=F}
par(mfrow=c(2, 2), mar=c(0,0.2,1,0.2))
g <- NULL
g = graph_from_adjacency_matrix(as.matrix(network.mat[[17]]),
                                mode='directed', weighted=TRUE,diag=F)
tiesSum = apply(g[], 1, sum)
# condition size based on # of ties
V(g)$size <- (tiesSum+0.5)*6
# only label if # ties greater than 10
V(g)$label <- ifelse( tiesSum>1, V(g)$name, NA )
V(g)$label.cex <- 0.6
plot(g,main=paste("Year:", names(network.mat)[17]),
     vertex.label=V(g)$label, 
     vertex.size=V(g)$size,
     edge.width=E(g)$weight,
     vertex.color =mypink, # change color of nodes
     vertex.label.color = "black", # change color of labels
     edge.curved=.25, # add a 25% curve to the edges
     edge.color=myblue, # change edge color to grey  
     layout=layout_with_fr,
     edge.arrow.size=0.2)
g <- NULL
g = graph_from_adjacency_matrix(as.matrix(network.mat[[18]]),
                                mode='directed', weighted=TRUE,diag=F)
tiesSum = apply(g[], 1, sum)
# condition size based on # of ties
V(g)$size <- tiesSum/2
# only label if # ties greater than 10
V(g)$label <- ifelse( tiesSum>1, V(g)$name, NA )
V(g)$label.cex <- 0.45
plot(g,main=names(network.mat)[18],
     vertex.label=V(g)$label, 
     vertex.size=V(g)$size,
     edge.width=E(g)$weight,
     vertex.color=mypink, # change color of nodes
     vertex.label.color = "black", # change color of labels
     edge.curved=.25, # add a 25% curve to the edges
     edge.color=myblue, # change edge color to grey  
     layout=layout_with_fr,
     edge.arrow.size=0.2)


```
\section{Exercise 2. Measurements \& Community detection}
\subsection{a}
We can measure influence with the degree of nodes. For example, we can count the total number of edges connected to a node to estimate the influence of an actor. By comparing network plot (above) and the degree (below), I am quite certain that I found an influential actor using degree. Additionally, I computed eigenvector centrality to find a node that is linked to other important nodes. When I compare actors who have the most degree and eigenvector centrality, they were almost the same. I report the most influential actor for each year and all years (in terms of the number of degree).

```{r degree, eval=T, echo=T, warning=F, message=F}
# to find an influential actor for each year

# create an empty list
influence.actor <- vector("list",18)
names(influence.actor) <- unique(nigeria$year)
eigen.actor <- vector("list",18)
names(eigen.actor) <- unique(nigeria$year)
# loop to find the maximum degree and eigenvector centrality
for(t in 1:17){
  temp <- graph.adjacency(as.matrix(network.mat[[t]]), mode='directed', weighted=TRUE,diag=F)
  degree <- igraph::degree(temp)
  eigen <- eigen_centrality(temp, directed=TRUE)$vector
  names(eigen) <- as.character(gsub("\n", " ", names(eigen)))
  names(degree) <- as.character(gsub("\n", " ", names(degree)))
  influence.actor[[t]] <- degree[which(degree==max(degree))]
  eigen.actor[[t]] <- eigen[which(eigen==max(eigen))]
}
# to find an influential actor overall
names(influence.actor)[18] <- "All Conflict"
names(eigen.actor)[18] <- "All Conflict"
temp <- graph.adjacency(as.matrix(network.mat[[18]]),mode='directed', weighted=TRUE,diag=F)
degree <- igraph::degree(temp)
eigen <- eigen_centrality(temp, directed=TRUE)$vector
names(eigen) <- as.character(gsub("\n", " ", names(eigen)))
names(degree) <- as.character(gsub("\n", " ", names(degree)))
influence.actor[[18]] <- degree[which(degree==max(degree))]
eigen.actor[[18]] <- eigen[which(eigen==max(eigen))]

influence.actor # for each year and all years
eigen.actor[[18]]
```
Police has the maximum degree for all years. Police also has the largest eigenvector centrality, which indicates that police is connected to other important actors.


\subsection{b}

Before I delve into the next question, I gauge data. Seeing a hierarchical clustering plot, it seems like there can be 3 (as little as) to 10 (as large as ) groups.
```{r gauge, eval=T, echo=T, warning=F, message=F}
library(sna)
g <- network(network.mat[[18]], directed=T)
eclusts <- equiv.clust(g)
plot(eclusts, hang=-1) 
```



```{r blockmodel, eval=T, echo=T, warning=F, message=F}



# running the block model using cross-validation

# save the node classification from each run

# out of sample cross validation to validate "k"

# report ROC statistics from each model

```


\section{Exercise 3. ERGMs}
```{r ergms, eval=T, echo=T, warning=F, message=F}
library(statnet)
set.seed(688346)
nigeria <- as.matrix(network.mat[[18]])
diag(nigeria) <- NA
nigeria <- as.network.matrix(nigeria)
nigeria %v% "group" <- ifelse(names(V(temp))=="Police (Nigeria)" | names(V(temp))=="Military (Nigeria)", 1, 0)
ergm.network <- ergm(nigeria ~ edges + mutual + nodematch('group'))
summary(ergm.network)
```

Here, my hypotheses is that a conflict in Nigeria would be reciprocal, and the government actors wouldn't involve in a conflict together. Since my hypotheses test reciprocity and group-homophily, I include "mutual" and "nodematch" variables in an ERGM model. The result shows that the number of ties, reciprocity, and the group-homophily variables are all statistically significant. The coefficient of mutual variable is positive and its effect size is large. This implies that reciprocity exists in Nigeria's conflict. The coefficient for nodematch is negative and significant as I expected. Government actors are less like to involve in a conflict with each other.




```{r ergmsdiag, eval=T, echo=T, warning=F, message=F}
mcmc.diagnostics(ergm.network)
ergm.network.gof = gof(ergm.network)
plot(ergm.network.gof)
```

To check the model fit, first, I examine the MCMC chains. They are well-mixed and stationary. Next, I check a goodness-of-fit of the model. Observed distributions (black lines) are generally within simulated ones. By looking at these plots, my model seems to be a reasonable approximation of the network. 